Image projections along an axis

### 2006-03-05 article/Article.tex
update timing and article
darcs-hash:20060305182553-2fc9d-a8da8ec3ed19cedb00b64321325af02ef22e40d4.gz

\begin{abstract}
\noindent
Image projection is a very common task in image analysis to reduce the dimension
of an image. A base filter is provided with some specialized filters which
implement different projection methods.
\end{abstract}

@@

Like \verb$AccumulateImageFilter$, \verb$ProjectionImageFilter$ can produce an
ouput image with the same dimension than the input image. In that case, all
the informations associated to the input image are copied to the output one,
and excepted:
\begin{itemize}
  \item the size of the image which is set to 1 on the projection dimension,
  \item the spacing of the image, which is modified on the projection dimension
to keep the physical size of the image on the output image,
  \item the index of the image wich is set to 0 on the projection dimension.
\end{itemize}
Basically, \verb$ProjectionImageFilter$ produce the same volume but with a
only 1 pixel on the projection dimension.

@@

However, we choose to move the dimensions and the indexes on those
dimensions, and so to not provide the ability to choose the direction.
The image is always projected with direction 1.
The user can still modify the output image with ITK filters like
\doxygen{FlipImageFilter}.

@@

The base class, \verb$ProjectionImageFilter$, is a re-write of
\verb$AccumulateImageFilter$, with some enhancement:
\begin{itemize}
  \item projection is delegated to a functor
  \item report its progress
  \item thread support
\end{itemize}
Different projection types are provided, with specialized classes:
\begin{itemize}
  \item maximum projection (\verb$MaximumProjectionImageFilter$)
  \item minimum projection (\verb$MinimumProjectionImageFilter$)
  \item mean projection (\verb$MeanProjectionImageFilter$)
  \item median projection (\verb$MedianProjectionImageFilter$)
  \item sum projection (\verb$SumProjectionImageFilter$)
  \item sigma projection (\verb$SigmaProjectionImageFilter$)
  \item binary projection (\verb$BinaryProjectionImageFilter$)
\end{itemize}
Other projections types can easily be created just by providing a new functor.

@@

A timing test comparing performance on a $371 \times 371 \times 34$
image showed that the new filter outperform the current implementation,
while it should have very similar performance.
This is the result of a problem in ITK 2.4.1, where the iterator with index
is a lot more efficient than the one without index. The new implementation
use the iterator with index for best performance, but really don't need
it, and should get even better performance with the iterator without
index once this problem will be fixed.

@@

\begin{table}[htbp]
\centering
\begin{tabular}{cc}
\hline
Filter & Execution time \\
\hline
\hline
\verb$MaximumProjectionImageFilter$ & 0.0841 s \\
\verb$MinimumProjectionImageFilter$  & 0.0846 s\\
\verb$SumProjectionImageFilter$  & 0.0872 s\\
\verb$MeanProjectionImageFilter$  & 0.0828 s\\
\verb$MedianProjectionImageFilter$  & 0.164 s\\
\verb$SigmaProjectionImageFilter$  & 0.123 s\\
\verb$BinaryProjectionImageFilter$  & 0.0848 s\\
\verb$AccumulateImageFilter$, sum  & 0.465 s\\
\verb$AccumulateImageFilter$, mean  & 0.465 s\\
\hline
\end{tabular}
\caption{Execution time with a single thread.\label{perf}}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics{results}
\caption{Execution time with an increasing number of threads.\label{results}}
\end{figure}

Figure~\ref{results} show an important increase of performance while using several
threads\footnote{It also show that the execution time is not really stable
and that the mean of 50 execution time is not enough to compare the execution
time with precision. This will be discussed in another article.}.

>>>>

\begin{abstract}
\noindent
Image projection is a very common task in image analysis to reduce the dimension
of an image. A base filter is provided with some specialized filters which
implement efficiently different projection methods.
\end{abstract}

@@

Like \verb$AccumulateImageFilter$, \verb$ProjectionImageFilter$ can produce an
ouput image with the same dimension than the input image. In that case, all
the informations associated to the input image are copied to the output one,
and excepted:
\begin{itemize}
  \item the size of the image which is set to 1 on the projection dimension,
  \item the spacing of the image, which is modified on the projection dimension
to keep the physical size of the image on the output image,
  \item the index of the image wich is set to 0 on the projection dimension.
\end{itemize}
Basically, \verb$ProjectionImageFilter$ produce the same volume but with
only 1 pixel on the projection dimension.

@@

However, we choose to move the dimensions {\em and} the indexes on those
dimensions, and so, to not provide the ability to choose the direction.
The image is always projected with direction 1.
The user can still modify the output image with ITK filters like
\doxygen{FlipImageFilter}.

@@

The base class, \verb$ProjectionImageFilter$, is a re-write of
\verb$AccumulateImageFilter$, with some enhancement:
\begin{itemize}
  \item projection is delegated to a functor,
  \item report its progress,
  \item thread support,
  \item dimension of ouput image can be smaller than dimension of the input image.
\end{itemize}
Different projection types are provided, with specialized classes:
\begin{itemize}
  \item maximum projection (\verb$MaximumProjectionImageFilter$)
  \item minimum projection (\verb$MinimumProjectionImageFilter$)
  \item mean projection (\verb$MeanProjectionImageFilter$)
  \item median projection (\verb$MedianProjectionImageFilter$)
  \item sum projection (\verb$SumProjectionImageFilter$)
  \item sigma projection (\verb$SigmaProjectionImageFilter$)
  \item binary projection (\verb$BinaryProjectionImageFilter$)
\end{itemize}
Other projections types can easily be created just by providing a new functor.

@@

A timing test comparing performance on a $371 \times 371 \times 34$
image showed that the new filter have better performance than the
\verb$AccumulateImageFilter$ for axis y and z, but worth for the x
axis. This is the result of the usage of a linear iterator in
\verb$ProjectionImageFilter$ and an image region iterator in
\verb$AccumulateImageFilter$.

@@

% #dim    size    max     min     sum     accSum  mean    accMean median  sigma   binary
% 0       371     0.02    0.02    0.0116  0.0128  0.0105  0.0128  0.0619  0.0504  0.019
% 1       371     0.0218  0.0218  0.0157  0.425   0.0153  0.424   0.0705  0.0538  0.021
% 2       34      0.0308  0.0308  0.0222  0.459   0.0223  0.46    0.104   0.0683  0.0299
\begin{table}[htbp]
\centering
\begin{tabular}{cccc}
\hline
Filter & Axis: x & Axis: y & Axis: z \\
\hline
\hline
\verb$MaximumProjectionImageFilter$  & 0.02 s   & 0.0218 s & 0.0308 s \\
\verb$MinimumProjectionImageFilter$  & 0.02 s   & 0.0218 s & 0.0308 s\\
\verb$MedianProjectionImageFilter$   & 0.0619 s & 0.0705 s & 0.104 s \\
\verb$SigmaProjectionImageFilter$    & 0.0504 s & 0.0538 s & 0.0683 s\\
\verb$BinaryProjectionImageFilter$   & 0.019  s & 0.021 s  & 0.0299 s\\
\verb$SumProjectionImageFilter$      & 0.0116 s & 0.0157 s & 0.0222 s\\
\verb$MeanProjectionImageFilter$     & 0.0105 s & 0.0153 s & 0.0223 s\\
\hline
\verb$AccumulateImageFilter$, sum    & 0.0128 s & 0.425 s  & 0.459 s \\
\verb$AccumulateImageFilter$, mean   & 0.0128 s & 0.424 s  & 0.46 s  \\
\hline
\end{tabular}
\caption{Execution time with a single thread.\label{perf}}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics{results}
\caption{Execution time with an increasing number of threads. Projection axis is z.\label{results}}
\end{figure}

% #th     max
% 1       0.0311
% 2       0.0197
% 3       0.017
% 4       0.0161
% 5       0.015
% 6       0.013
% 7       0.0119
% 8       0.0108
% 9       0.0116
% 10      0.0115
\begin{table}[htbp]
\centering
\begin{tabular}{ccc}
\hline
Number of threads & Execution time & Speedup \\
\hline
\hline
1      & 0.0311 s & 1 \\
2      & 0.0197 s & 1.58\\
3      & 0.017  s & 1.83\\
4      & 0.0161 s & 1.93\\
5      & 0.015  s & 2.07\\
6      & 0.013  s & 2.39\\
7      & 0.0119 s & 2.61\\
8      & 0.0108 s & 2.88\\
9      & 0.0116 s & 2.68\\
10     & 0.0115 s & 2.7\\
\hline
\end{tabular}
\caption{Execution time and speedup with an increasing number of threads.\label{thread}}
\end{table}

Figure~\ref{results} and Table~\ref{thread} show an important increase of performance while using several
threads.
